Accelerating Burn Speed of Birch Wood Popsicle Sticks

### 2022-05-02 paper/paper.Rmd
Add inline citations and update factorial design, data table

This gives us a $4!$ experimental design, since we have four treatment effects and no other factors.

@@

@@

Data were cleaned using Python version 3.7 (Python, Wilmington, DE) and Pandas version 1.0.3. Data were analyzed using R version 4.0.3 (The R Project for Statistical Computing, Indianapolis, IN). The Skillings-Mack procedure was run in R using the NSM3 package, version 1.16. The Wilcoxon rank-sum tests were run using the stats package in R.

@@

We attempted to fix the data shape using a data transformation, including the $y^2$, $y^3$, $\sqrt{y}$, and $\ln{y}$ transformations. However, none of these transformations changed the data enough to change the dispersion or the shape of the distribution to be more normal. Instead, we chose to use a non-parametric method instead. For the type of data, we chose to use the Skillings-Mack procedure. The Skillings-Mack procedure is a distribution free test for general alternatives for data from an arbitrary incomplete block design. This means that it can be used where there are missing observations in a single block. Its parametric cousin is the two-way ANOVA, and it can be used as a proxy for an overall F-test where the conditions are not met and the data are from an arbitrary incomplete block design.

@@

$$
H_A: [\alpha_\text{acetone}, \alpha_\text{dry}, \alpha_\text{ethanol}, \alpha_\text{isopropyl} \text{ not all equal}],
$$

when $c_{ij} \equiv 1$, for $i = 1, 2, 3, 4$ and $j = 1, 2, 3, 4, 5, 6$.

@@

1. Run a Wilcoxon rank-sum test on the data for all $6 = {4\choose2}$ treatment pairwise comparisons, where $H_0: F(t) = G(t)$ and $H_A: F(t) \neq G(t)$, where $F(t)$ represents the cumulative distribution function for one treatment, and $G(t)$ represents the cumulative distribution function for another treatment
2. Keep the block and response values together, but randomly assign a new treatment for each observation, then calculate the Wilcoxon rank sum for all $6 = {4\choose2}$ treatment pairwise comparisons. This will spread the block effect across all treatments. Repeat this $r$ times, and store each of the test statistics for the pairwise comparisons for each of these tests.
3. For each of the six pairwise comparisons groups, compare the Bonferroni-adjusted \nth{5} and \nth{95} percentiles of the $r$ simulated random-treatment Wilcoxon rank-sum test statistics against the test statistic from each of the pairwise Wilcoxon tests from the actual data. If the actual Wilcoxon rank-sum test statistic falls inside of the Bonferroni-adjusted (0.05, 0.95) percentile from the random-treatment distribution, there is a significant difference between the two groups.

@@

In Table 1, we found that the simulated $U$'s are not statistically different from the actual $U$'s for the acetone-dry, acetone-isopropyl, and isopropyl-dry combinations. This means that there is no statistically significant difference between the results for isopropyl alcohol, acetone, and no accelerant.

@@

Our goal is to maximize the amount burnt, so the metric for success is to minimize the unburnt distance. We saw in the boxplot from Figure 5 that the best way to decrease unburnt distance is to use acetone. Table 1 tells us that there was no significant difference between acetone, dry, or isopropyl alcohol. Therefore, we recommend that to maximize the amount burnt, to soak birch popsicle sticks in acetone, isopropyl alcohol, or no accelerant at all.

@@

>>>>

This gives us a one-factor experimental design with four levels and a block.

@@

The treatments for each block are given in Table 1.

| Block | Treatments                             |
|-------|----------------------------------------|
| 1     | isopropyl, acetone, ethanol, acetone   |
| 2     | ethanol, dry, acetone, ethanol         |
| 3     | ethanol, isopropyl, isopropyl, dry     |
| 4     | isopropyl, ethanol, acetone, isopropyl |
| 5     | ethanol, acetone, dry, acetone         |
| 6     | dry, isopropyl, dry, dry               |

Table: Blocks and Treatments Table

@@

Data were cleaned using Python version 3.7 (Python, Wilmington, DE) and Pandas version 1.0.3. [@vanrossum2009; @pandas2020] Data were analyzed using R version 4.0.3 (The R Project for Statistical Computing, Indianapolis, IN). [@R-base] The Skillings-Mack procedure was run in R using the NSM3 package, version 1.16. [@schneider2021] The Wilcoxon rank-sum tests were run using the stats package in R.

@@

We attempted to fix the data shape using a data transformation, including the $y^2$, $y^3$, $\sqrt{y}$, and $\ln{y}$ transformations. However, none of these transformations changed the data enough to change the dispersion or the shape of the distribution to be more normal. Instead, we chose to use a non-parametric method instead. For the type of data, we chose to use the Skillings-Mack procedure. [@hollander2013] The Skillings-Mack procedure is a distribution free test for general alternatives for data from an arbitrary incomplete block design. This means that it can be used where there are missing observations in a single block. Its parametric cousin is the two-way ANOVA, and it can be used as a proxy for an overall F-test where the conditions are not met and the data are from an arbitrary incomplete block design.

@@

and

$$
H_A: [\alpha_\text{acetone}, \alpha_\text{dry}, \alpha_\text{ethanol}, \alpha_\text{isopropyl} \text{ not all equal}].
$$

@@

1. Run a Wilcoxon rank-sum test on the data for all $6 = {4\choose2}$ treatment pairwise comparisons, where $H_0: F(t) = G(t)$ and $H_A: F(t) \neq G(t)$, where $F(t)$ represents the cumulative distribution function for one treatment, and $G(t)$ represents the cumulative distribution function for another treatment [@hollander2013]
2. Keep the block and response values together, but randomly assign a new treatment for each observation, then calculate the Wilcoxon rank sum for all $6 = {4\choose2}$ treatment pairwise comparisons. This will spread the block effect across all treatments. Repeat this $r$ times, and store each of the test statistics for the pairwise comparisons for each of these tests.
3. For each of the six pairwise comparisons groups, compare the Bonferroni-adjusted \nth{5} and \nth{95} percentiles of the $r$ simulated random-treatment Wilcoxon rank-sum test statistics against the test statistic from each of the pairwise Wilcoxon tests from the actual data. If the actual Wilcoxon rank-sum test statistic falls inside of the Bonferroni-adjusted (0.05, 0.95) percentile from the random-treatment distribution, there is a significant difference between the two groups.

@@

In Table 2, we found that the simulated $U$'s are not statistically different from the actual $U$'s for the acetone-dry, acetone-isopropyl, and isopropyl-dry combinations. This means that there is no statistically significant difference between the results for isopropyl alcohol, acetone, and no accelerant.

@@

Our goal is to maximize the amount burnt, so the metric for success is to minimize the unburnt distance. We saw in the boxplot from Figure 5 that the best way to decrease unburnt distance is to use acetone. Table 2 tells us that there was no significant difference between acetone, dry, or isopropyl alcohol. Therefore, we recommend that to maximize the amount burnt, to soak birch popsicle sticks in acetone, isopropyl alcohol, or no accelerant at all.

@@

## Availability

The code and data are open-source and are available at \href{https://github.com/kim3-sudo/popsiclefire}{https://github.com/kim3-sudo/popsiclefire}.
